Storm surges in the Strait of Georgia simulated with a regional model

### 2015-01-03 AO_guides/manuscriptAO.tex
Keith't typographical comments

%my attempt at physical oceanography in SoG- Susan feel free to edit.
%Several features related to the physical oceanography of this region render it a challenge to model numerically. The outflow from the Fraser River in the late spring is the primary source of fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months \citep{thomson1981oceanography}. Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia. It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation \citep{masson2004modelling}. The strong and fast tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, set a fairly restrictive time step and impose a particular challenge for numerical models. Yet, several modelling efforts for this region exist, ranging in choices of grid structure, mixing parametrizations, and domain size and extent.

@@

More recently, modelling studies have focused on baroclinic features. \citet{masson2004modelling} used the Princeton Ocean Model (POM) in a regional mode to examine the seasonal variability of deep water properties in the Strait of Georgia noting a strong sensitivity to vertical mixing over the sills. This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography. A finite volume model on an unstructured grid by \citet{yang2010multi} involved a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary. \citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia and noted that their semi-diurnal amplitudes were generally too low. An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}. Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.  Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling this region.

@@

River input provides a significant volume of fresh water to the Salish Sea and can influence stratification, circulation, and primary productivity. However, most rivers in the domain are not gauged so parametrizations were required to represent river flow. \citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation. Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages. Freshwater runoff from each watershed was divided between the rivers in that watershed. The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website. The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%). The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperature. A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.

@@

Model results have been compared with observations by calculating both the total modelled water level and the modelled residual at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}). The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers. Water level observations are taken from the Canadian Tides and Water Level Data Archive \citep{DFOObservations} and are used to calculate the observed residual, defined as the observed water level minus tidal predictions. Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior to the surge event of interest.

@@

In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}). The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in sync with observations. The surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide. As such, it is of minor consequence that the timing of the observed and modelled maximum residual do not match perfectly.

@@

Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}). Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly. Second, a simulation with all forcing except the atmospheric forcing. Third, a simulation with all forcing except the sea surface height anomaly. Fourth, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure. Comparisons between these simulations indicate the relative importance the the open boundary forcing, the local wind forcing, and the local atmospheric pressure forcing. Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction. Contributions due to stratification and bathymetric resolution will be left to a future study.

@@

These comparisons point to the importance of including, as a forcing condition at the open boundary, the surge entering the domain from the Pacific Ocean. Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca. In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition. Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.

@@

Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}). The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.  In contrast, the tide-surge interaction at Victoria is hardly noticeable. Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands. The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation. However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge. So, in operational settings, performing a surge-only simulation and adding the predicted tides \textit{a posteriori} could be more computationally efficient with only a small decrease in the accuracy of the prediction.

@@

\begin{table}[h]
\centering 
\tbl{Comparisons of modelled and observed amplitudes and phases. $R$ is the ratio of modelled to observed amplitude. $\Delta \phi$ is the difference between modelled and observed phase. $D$ is the complex difference. The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity. }
{\begin{tabular}{l l l l c c c c c c} 
\hline 
& & & & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
\hline 
            & Station name     & Longitude & Latitude & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
\hline 
1 & Port Renfrew & -124.43 & 48.55 & 1.16 & -9.4 & 16.63  & 1.03 & -4.94 & 4.17 \\
2 & Sheringham Point & -123.92 & 48.38 & 1.19 & -16.47 & 17.95  & 1 & -4.44 & 4.24 \\
3 & Pedder Bay & -123.55 & 48.33 & 0.96 & -27.39 & 15.91  & 0.99 & -2.41 & 2.66 \\
4 & Esquimalt & -123.43 & 48.43 & 0.91 & -23.09 & 14.40  & 0.99 & -1.04 & 1.21 \\
5 & Victoria & -123.37 & 48.42 & 0.89 & -17.77 & 11.62  & 1.04 & -1.83 & 3.22 \\
6 & Clover Point & -123.35 & 48.41 & 0.83 & -17.95 & 13.39  & 1.03 & -2.23 & 3.14 \\
7 & Finnerty Cove & -123.3 & 48.47 & 0.83 & -8.24 & 9.47  & 1.02 & -2.19 & 3.25 \\
8 & Fulford Harbour & -123.45 & 48.77 & 0.82 & -3.08 & 10.78  & 1 & -0.63 & 0.83 \\
9 & Tumbo Channel & -123.11 & 48.79 & 1.05 & 2.03 & 4.43  & 1.04 & 0.66 & 3.34 \\
10 & Patos Island & -122.97 & 48.78 & 1.03 & 0.26 & 2.32  & 1.04 & -1.5 & 3.92 \\
11 & Whaler Bay & -123.33 & 48.89 & 1 & -0.76 & 1.10  & 1.01 & -0.48 & 1.19 \\
12 & Tsawwassen & -123.13 & 48.99 & 1 & 0.92 & 1.31  & 1.01 & 0.19 & 1.09 \\
13 & Sandheads & -123.3 & 49.1 & 1.02 & -1.05 & 2.58  & 1.04 & -0.59 & 3.15 \\
14 & Point Grey & -123.27 & 49.25 & 0.98 & -3.31 & 5.83  & 0.96 & -0.68 & 3.63 \\
15 & Point Atkinson & -123.25 & 49.33 & 1.01 & -0.48 & 1.07  & 1 & -0.56 & 0.94 \\
16 & Gibsons Landing & -123.5 & 49.4 & 0.98 & 0.75 & 2.34  & 1 & 0.96 & 1.47 \\
17 & Winchelsea & -124.08 & 49.3 & 1.01 & -0.19 & 0.78  & 1.01 & 0.51 & 1.30 \\
18 & Halfmoon Bay & -123.92 & 49.52 & 1.01 & -0.15 & 0.83  & 1.01 & 0.15 & 0.72\\ 
19 & Irvines Landing & -124.05 & 49.63 & 1 & -0.32 & 0.59  & 1.01 & 0.24 & 1.12 \\
20 & Powell River & -124.55 & 49.86 & 1 & -1.72 & 3.05  & 1 & -0.01 & 0.43 \\
21 & Little River & -124.92 & 49.74 & 1.03 & 0.07 & 2.54  & 1.01 & 0.33 & 0.73\\ 
22 & Lund & -124.76 & 49.98 & 1.02 & -2.52 & 4.98  & 1.02 & -0.95 & 2.45 \\
23 & Twin Islets & -124.93 & 50.03 & 1.02 & -2.05 & 4.22  & 1.01 & -0.24 & 0.72\\ 
24 & Campbell River & -125.24 & 50.04 & 0.99 & 0.51 & 1.12  & 1 & 0.94 & 1.39 \\
25 & Maude Island E & -125.33 & 50.13 & 1.13 & -10.02 & 12.62  & 0.97 & -1.44 & 2.89\\ 
26 & Nymphe Cove & -125.36 & 50.13 & 1.02 & -5.43 & 6.04  & 0.97 & -0.29 & 2.02 \\
27 & Seymour Narrows & -125.34 & 50.13 & 0.63 & 31.31 & 53.40  & 1.14 & 10.91 & 17.28 \\
28 & Brown Bay & -125.37 & 50.16 & 0.98 & -5.63 & 9.26  & 0.98 & -0.04 & 1.17 \\
29 & Chatham Point & -125.44 & 50.33 & 1.09 & -5.49 & 12.18  & 0.98 & -3.18 & 3.85 \\
30 & Kelsey Bay & -125.96 & 50.39 & 1 & -1.69 & 3.50  & 0.98 & -0.39 & 1.00 \\
31 & Yorke Island & -125.97 & 50.44 & 0.99 & 1.59 & 3.32  & 1.01 & 1.23 & 1.37\\ 
\hline 
&   Mean Error (cm/deg) & & & 4.23  & 6.5 & 8.05   & 1.56 & 1.49 & 2.58 \\
&   RMS Error (cm/deg)  & & & 7.67  & 10.73 & 12.73 & 2.41 & 2.57 & 3.91 \\
\hline  
\end{tabular}}
\label{tab:comparison} 
\end{table}

>>>>

%my attempt at physical oceanography in SoG- Susan feel free to edit.
%Several features related to the physical oceanography of this region render it a challenge to model numerically. The outflow from the Fraser River in the late spring is the primary source of fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months \citep{thomson1981oceanography}. Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia. It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation \citep{masson2004modelling}. The strong tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, set a fairly restrictive time step and impose a particular challenge for numerical models. Several modelling efforts for this region exist, varying in choices of grid structure, mixing parametrizations, and domain.

@@

More recently, modelling studies have focused on baroclinic features. \citet{masson2004modelling} used the Princeton Ocean Model (POM) in a regional mode to examine the seasonal variability of deep water properties in the Strait of Georgia and noted a strong sensitivity to vertical mixing over the sills. This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography. 
%A finite volume model on an unstructured grid by \citet{yang2010multi} involved a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary. 
\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia. They noted that their semi-diurnal amplitudes were generally too low. An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}. Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.  Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling this region.

@@

River input provides a significant volume of fresh water to the Salish Sea and can influence stratification, circulation, and primary productivity. However, most rivers in the domain are not gauged so parametrizations were required to represent river flow. \citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation. Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages. Freshwater runoff from each watershed was divided among the rivers in that watershed. The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website. The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%). The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperature. A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.

@@

Model results have been compared with observations by calculating both the total modelled water level and the modelled residual at four permanent tidal measuring stations: Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}). The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers. Water level observations are taken from the Canadian Tides and Water Level Data Archive \citep{DFOObservations} and are used to calculate the observed residual, defined as the observed water level minus tidal predictions. Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior to the surge event of interest.

@@

In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}). The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in phase with observations. The surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide. As such, it is of minor consequence that the timing of the observed and modelled maximum residual do not match perfectly.

@@

Next, an assessment of the factors that are most important in the storm surge forcing conditions is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}). Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly. Second, a simulation with all forcing except the atmospheric forcing. Third, a simulation with all forcing except the sea surface height anomaly. Fourth, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure. Comparisons between these simulations indicate the relative importance the the open boundary forcing, the local wind forcing, and the local atmospheric pressure forcing. Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction. Contributions due to stratification and bathymetric resolution will be left to a future study.

@@

These comparisons point to the importance of including, as a forcing condition at the open boundary, the surge entering the domain from the Pacific Ocean. Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca. In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition. Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain, consistent with the results of the present study.

@@

Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}). The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.  In contrast, the tide-surge interaction at Victoria is hardly noticeable. Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands. The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation. However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge. So, in operational settings, performing a surge-only simulation and adding the predicted tides \textit{a posteriori} could be more computationally efficient with only a small decrease in the accuracy of the prediction.

@@

\begin{table}[h]
\centering 
\tbl{Comparisons of modelled and observed amplitudes and phases. $R$ is the ratio of modelled to observed amplitude. $\Delta \phi$ is the difference between modelled and observed phase. $D$ is amplitude of the complex difference. The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity. }
{\begin{tabular}{l l l l c c c c c c} 
\hline 
& & & & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
\hline 
            & Station name     & Longitude & Latitude & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
\hline 
1 & Port Renfrew & -124.43 & 48.55 & 1.16 & -9.4 & 16.63  & 1.03 & -4.94 & 4.17 \\
2 & Sheringham Point & -123.92 & 48.38 & 1.19 & -16.47 & 17.95  & 1 & -4.44 & 4.24 \\
3 & Pedder Bay & -123.55 & 48.33 & 0.96 & -27.39 & 15.91  & 0.99 & -2.41 & 2.66 \\
4 & Esquimalt & -123.43 & 48.43 & 0.91 & -23.09 & 14.40  & 0.99 & -1.04 & 1.21 \\
5 & Victoria & -123.37 & 48.42 & 0.89 & -17.77 & 11.62  & 1.04 & -1.83 & 3.22 \\
6 & Clover Point & -123.35 & 48.41 & 0.83 & -17.95 & 13.39  & 1.03 & -2.23 & 3.14 \\
7 & Finnerty Cove & -123.3 & 48.47 & 0.83 & -8.24 & 9.47  & 1.02 & -2.19 & 3.25 \\
8 & Fulford Harbour & -123.45 & 48.77 & 0.82 & -3.08 & 10.78  & 1 & -0.63 & 0.83 \\
9 & Tumbo Channel & -123.11 & 48.79 & 1.05 & 2.03 & 4.43  & 1.04 & 0.66 & 3.34 \\
10 & Patos Island & -122.97 & 48.78 & 1.03 & 0.26 & 2.32  & 1.04 & -1.5 & 3.92 \\
11 & Whaler Bay & -123.33 & 48.89 & 1 & -0.76 & 1.10  & 1.01 & -0.48 & 1.19 \\
12 & Tsawwassen & -123.13 & 48.99 & 1 & 0.92 & 1.31  & 1.01 & 0.19 & 1.09 \\
13 & Sandheads & -123.3 & 49.1 & 1.02 & -1.05 & 2.58  & 1.04 & -0.59 & 3.15 \\
14 & Point Grey & -123.27 & 49.25 & 0.98 & -3.31 & 5.83  & 0.96 & -0.68 & 3.63 \\
15 & Point Atkinson & -123.25 & 49.33 & 1.01 & -0.48 & 1.07  & 1 & -0.56 & 0.94 \\
16 & Gibsons Landing & -123.5 & 49.4 & 0.98 & 0.75 & 2.34  & 1 & 0.96 & 1.47 \\
17 & Winchelsea & -124.08 & 49.3 & 1.01 & -0.19 & 0.78  & 1.01 & 0.51 & 1.30 \\
18 & Halfmoon Bay & -123.92 & 49.52 & 1.01 & -0.15 & 0.83  & 1.01 & 0.15 & 0.72\\ 
19 & Irvines Landing & -124.05 & 49.63 & 1 & -0.32 & 0.59  & 1.01 & 0.24 & 1.12 \\
20 & Powell River & -124.55 & 49.86 & 1 & -1.72 & 3.05  & 1 & -0.01 & 0.43 \\
21 & Little River & -124.92 & 49.74 & 1.03 & 0.07 & 2.54  & 1.01 & 0.33 & 0.73\\ 
22 & Lund & -124.76 & 49.98 & 1.02 & -2.52 & 4.98  & 1.02 & -0.95 & 2.45 \\
23 & Twin Islets & -124.93 & 50.03 & 1.02 & -2.05 & 4.22  & 1.01 & -0.24 & 0.72\\ 
24 & Campbell River & -125.24 & 50.04 & 0.99 & 0.51 & 1.12  & 1 & 0.94 & 1.39 \\
25 & Maude Island E & -125.33 & 50.13 & 1.13 & -10.02 & 12.62  & 0.97 & -1.44 & 2.89\\ 
26 & Nymphe Cove & -125.36 & 50.13 & 1.02 & -5.43 & 6.04  & 0.97 & -0.29 & 2.02 \\
27 & Seymour Narrows & -125.34 & 50.13 & 0.63 & 31.31 & 53.40  & 1.14 & 10.91 & 17.28 \\
28 & Brown Bay & -125.37 & 50.16 & 0.98 & -5.63 & 9.26  & 0.98 & -0.04 & 1.17 \\
29 & Chatham Point & -125.44 & 50.33 & 1.09 & -5.49 & 12.18  & 0.98 & -3.18 & 3.85 \\
30 & Kelsey Bay & -125.96 & 50.39 & 1 & -1.69 & 3.50  & 0.98 & -0.39 & 1.00 \\
31 & Yorke Island & -125.97 & 50.44 & 0.99 & 1.59 & 3.32  & 1.01 & 1.23 & 1.37\\ 
\hline 
&   Mean Error (cm/deg) & & & 4.23  & 6.5 & 8.05   & 1.56 & 1.49 & 2.58 \\
&   RMS Error (cm/deg)  & & & 7.67  & 10.73 & 12.73 & 2.41 & 2.57 & 3.91 \\
\hline  
\end{tabular}}
\label{tab:comparison} 
\end{table}
